---
title: "Injury"
subtitle: "AHPS Injury Project"
date: "2024-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries and Datasets
```{r}
library(dplyr) # Data Manipulation Package %>% 
library(tidyr) # Data Cleaning Package
library(readxl) # Excel
library(tidyverse) # Cleaning/Manipulating
library(purrr) # Initializing Lists to Data Frames
library(lubridate) # Working with Dates
library(openxlsx) # Excel 
library(stringr) # String Splitting
library(ggplot2) # Plotting
library(stringr) # Prefix/Suffix Editting
library(manipulate) # Easy Chart Manipulation
library(shiny) # Interactive Charting
library(survival) # CoxPH Model
library(gridExtra) # Plotting Charts Together
library(grid) # Grid Manipulation 
```

## Load Dataframes
```{r}


# List all Files
list_excel_files <- list.files(path = "Path", full.name = TRUE)

# Function for looping through all files
read_excel_file <- function(file) {
  if(is.na(file)) stop("No file path") # Testing if Path exists
df <- readxl::read_excel(file, skip = 7, col_names = TRUE)
df$`Body Weight [kg]` <- as.numeric(df$`Body Weight [kg]`)
df
}

# Create Initial Force plate Data frame
fp <- purrr::map_dfr(.x = list_excel_files, .f = read_excel_file)

# Correctly Format Names
split_name <- strsplit(fp$Athlete, " ")

fp$Athlete <- sapply(split_name, function(name) {
  paste(rev(name), collapse = ", ")
})

# Convert the datetime column to a Date object
fp$`Test Date` <- as.Date(fp$`Test Date`, format = "%Y-%m-%d")

# Remove unneeded files
rm(split_name)

# Combine Athlete Standing Weights and Body Weight variables
fp <- fp %>% 
  mutate(`Body Weight [kg]` = coalesce(`Body Weight [kg]`, `Athlete Standing Weight [kg]`))

# Remove Athlete Standing Weight
fp$`Athlete Standing Weight [kg]` <- NULL

# Export FP data frame
write.xlsx(fp, "formatted_FP.xlsx", rowNames = FALSE)

injury_df <- read_excel("Path")
combined_fp_injury <- read_excel("Path")

# Resorting Data frame
combined_fp_injury <- combined_fp_injury %>%
  select(incident_date, `Test Date`, DaysDifference, athlete_name, Athlete, everything())

```

## Formatting and Tweaking Injury Data Frame
```{r}
# Format some Injury Data
# Fix one observation
injury_df <- injury_df %>%
  mutate(incident_sport = if_else(incident_sport == "Other - Sports", "American Football", incident_sport))



# Separate Mens/Womens BBall

# Function to extract sports information
extract_sports <- function(input_string) {
  if (grepl("Womens Basketball", input_string, ignore.case = TRUE)) {
    return("Womens Basketball")
  } else if (grepl("Mens Basketball", input_string, ignore.case = TRUE)) {
    return("Mens Basketball")
  } else {
    return(NA)
  }
}

# Use dplyr to apply the function to the data frame
injury_df <- injury_df %>%
  mutate(sports_category = sapply(sport_identity, extract_sports))


injury_df <- injury_df %>% 
  mutate(incident_sport = ifelse(incident_sport == "Basketball", sports_category, incident_sport))

injury_df$sports_category <- NULL


# Separate Injuries from Left / Right

extract_words_between_brackets <- function(input_string) {
  
  # Define a regular expression pattern to match words between square brackets
  pattern <- "\\[([^\\]]+)\\]"

  # Use str_extract to find the first match of the pattern in the input string
  match <- str_extract(input_string, pattern)

  # If a match is found, extract the words; otherwise, return NA
  if (!is.na(match)) {
    words <- str_extract(match, "\\b\\w+\\b")
    return(ifelse(!is.na(words), words, NA))
  } else {
    return(NA)
  }
}


# 57 observations do not have Right/Left/Bilateral !!!!!!!!!!!!!!!
injury_df <- injury_df %>%
  mutate(area_of_body = sapply(issue_description, extract_words_between_brackets))

# sum(is.na(injury_df$area_of_body))


write.xlsx(injury_df, "injury_symmetry_cleaned.xlsx", rowNames = FALSE)


```

## Exploratory

## Correlation and Regression Analysis on FP Metrics
```{r}

# List of metric names (replace with your actual variable names)
metric_names <- names(combined_fp_injury)[28:ncol(combined_fp_injury)]

# Remove Columns with N/A observations
columns_with_na <- colnames(combined_fp_injury)[colSums(is.na(combined_fp_injury)) > 0]
na_counts <- colSums(is.na(combined_fp_injury[, columns_with_na]))

# Remove Columns with > 500 NA
columns_to_remove <- columns_with_na[na_counts > 500]

combined_fp_injury <- combined_fp_injury %>% 
  select(-columns_to_remove)

# Check Work
new_columns_with_na <- colnames(combined_fp_injury)[colSums(is.na(combined_fp_injury)) > 0]

# Set Up Data Frame for Slopes
slopes_df <- data.frame(
  Metric = character(),
  Slope = numeric()
)

# Loop through each metric and calculate the slope
for (metric in metric_names) {
  model <- lm(combined_fp_injury[[metric]] ~ days_before_injury, data = combined_fp_injury)
  slope <- coef(model)[2]  # Extract the slope
  slopes_df <- bind_rows(slopes_df, data.frame(Metric = metric, Slope = slope))
}

correlation_df <- data.frame(
  Metric = character(),
  Correlation = numeric()
)

# Loop through each metric and calculate the correlation coefficient
for (metric in metric_names) {
  correlation <- cor(combined_fp_injury[[metric]], combined_fp_injury$days_before_injury)
  correlation_df <- bind_rows(correlation_df, data.frame(Metric = metric, Correlation = correlation))
}
  
# Looking at these relationships visually
ggplot(data = combined_fp_injury, aes(x = days_before_injury, y = `Start of Integration [s]`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  scale_x_reverse()


```

## Initial Analysis (Using Mens Basketball)
```{r}

# Filter to just MBB observations
mbb_combined <- combined_fp_injury %>% 
  filter(incident_sport == "Mens Basketball")

# Select Relevant Variables
mbb_combined <- mbb_combined %>% 
  select(Athlete, `Test Date`, incident_date, DaysDifference, body_area, issue_description, area_of_body, 
         `Body Weight [kg]`, `Jump Height (Flight Time) [cm]`, `Flight Time [ms]`, `RSI-modified [m/s]`, 
         `Flight Time:Contraction Time`, `Eccentric Peak Force Asymmetry [% L,R]`, 
         `Eccentric Deceleration RFD Asymmetry [% L,R]`, `Eccentric Deceleration RFD [N/s]`, 
         `Eccentric Acceleration Phase Duration [s]`, `Braking Phase Duration [s]`)



# Create Baseline
# Compare Month prior to Baselines

# [% L,R] - if L bigger + L smaller

# Separating Data frames into Baseline and Comparisons
baseline_mbb <- mbb_combined %>% 
  filter(DaysDifference > 30)

month_prior_mbb <- mbb_combined %>% 
  filter(DaysDifference <= 30)

# Adds Identifier to Force Plate Metric Columns
baseline_mbb <- baseline_mbb %>% 
  rename_at(vars(8:ncol(baseline_mbb)), ~paste("var_", ., sep = ""))

# Selects median observation for each incident by athlete
median_by_injury_observation <- baseline_mbb %>%
  group_by(Athlete, incident_date) %>%
  summarise(across(starts_with("var_"), median))

# Removes Identifier
baseline_mbb <- baseline_mbb %>% 
  rename_at(vars(starts_with("var_")), ~str_remove(., "var_"))

median_by_injury_observation <- median_by_injury_observation %>% 
  rename_at(vars(starts_with("var_")), ~str_remove(., "var_"))

# Athlete Specific 
median_single_athlete <- median_by_injury_observation %>% 
  filter(Athlete == "Athlete_1")

month_before_single_athlete <- month_prior_mbb %>% 
  filter(Athlete == "Athlete_1")

# Remove Dates that do not match between data frames
month_before_single_athlete <- month_before_single_athlete %>% 
  filter(incident_date %in% median_single_athlete$incident_date)

median_single_athlete <- median_single_athlete %>% 
  filter(incident_date %in% month_before_single_athlete$incident_date)

# Identifier for Median
median_single_athlete <- median_single_athlete %>% 
  rename_at(vars(3:ncol(median_single_athlete)), ~paste("med_", ., sep = ""))

# Join the data frames
combined_single_athlete <- merge(month_before_single_athlete, median_single_athlete, by = c("Athlete", "incident_date"))

# Finds value of Median - Jump
combined_single_athlete <- combined_single_athlete %>%
  mutate(across(starts_with("med_"), 
                list(dif = ~. - get(sub("med_", "", cur_column()))), .names = "dif_{.col}"))

# Format it correctly
combined_single_athlete <- combined_single_athlete %>%
  mutate(across(starts_with("dif_"), ~ . * -1))
                

ggplot(data = combined_single_athlete, aes(x = DaysDifference, y = `dif_med_Eccentric Acceleration Phase Duration [s]`)) +
  geom_point() +
  geom_line(y = 0, col = "red") +
  labs(x = "Days Until Injury",
       y = "Eccentric Acceleration Phase Duration [s]") +
  scale_x_reverse() +
  theme_minimal()

# Get Column Names that start with "dif_"
grep("dif_", colnames(combined_single_athlete), value = TRUE)





```

## Applying this Analysis to whole team

```{r}

# Join the data frames
combined_mbb <- merge(month_prior_mbb, median_by_injury_observation, by = c("Athlete", "incident_date"), all.x = TRUE)

# Drop Values without a Baseline Value
combined_mbb <- combined_mbb %>%
filter(!is.na(`med_Body Weight [kg]`))

# Display the number of NA values in each column
na_counts <- colSums(is.na(combined_mbb))
print(na_counts)

# Finds value of Median - Jump
combined_mbb <- combined_mbb %>%
  mutate(across(starts_with("med_"),
  list(dif = ~. - get(sub("med_", "", cur_column()))), .names = "dif_{.col}"))

# Format it correctly
combined_mbb <- combined_mbb %>%
  mutate(across(starts_with("dif_"), ~ . * -1))



# Plotted each observation by Athlete and Incident Date
ggplot(data = combined_mbb, aes(x = DaysDifference, y = `dif_med_Jump Height (Flight Time) [cm]`, 
                                group = interaction(Athlete, incident_date), color = Athlete)) +
  geom_line() +
  geom_point() +
  geom_line(y = 0, col = "red") +
  labs(x = "Days Until Injury") +
  scale_x_reverse() +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(~Athlete + incident_date, scales = "free", ncol = 10)

# Plotted separate lines for each injury for each Athlete
ggplot(data = combined_mbb, aes(x = DaysDifference, y = `dif_med_Jump Height (Flight Time) [cm]`, 
                                group = interaction(Athlete, incident_date), color = incident_date)) +
  geom_line() +
  geom_point() +
  geom_line(y = 0, col = "red") +
  labs(x = "Days Until Injury") +
  scale_x_reverse() +
  theme_minimal() +
  guides(color = FALSE) +
  facet_wrap(~Athlete, scales = "free")




#  ("dif_med_Body Weight [kg]", "dif_med_Jump Height (Flight Time) [cm]", "dif_med_Flight Time [ms]", "dif_med_RSI-modified [m/s]", 
#                     "dif_med_Eccentric Peak Force Asymmetry [% L,R]", "dif_med_Flight Time:Contraction Time", "dif_med_Eccentric Deceleration RFD [N/s]", 
#                     "dif_med_Eccentric Deceleration RFD Asymmetry [% L,R]", "dif_med_Eccentric Acceleration Phase Duration [s]", 
#                     "dif_med_Braking Phase Duration[s]")


grep("dif_", colnames(combined_mbb), value = TRUE)

```

## R Shiny Graph for Just Mens Basketball

```{r}
library(shiny)

ui <- fluidPage(
  selectInput("variable", "Choose Y-axis variable", choices = colnames(combined_mbb)[28:37]),
  plotOutput("scatterPlot")
)

server <- function(input, output) {
  output$scatterPlot <- renderPlot({
    ggplot(data = combined_mbb, aes_string(x = "DaysDifference", y = paste0("`", input$variable, "`"))) +
      geom_point() +
      geom_smooth(method = "lm", se = TRUE, color = "blue") +
      geom_line(y = 0, col = "red") +
      labs(title = paste("Scatter Plot with", input$variable, "on Y-axis")) +
      theme_minimal() +
      scale_x_reverse()
  })
}

shinyApp(ui, server)


```

## Build out R Shiny Application that allows to switch between sports

## Want to filter by Sport, Variable, Injury Type
## Where Median/Baseline is Created
```{r}

# Filter to relevant metrics
filtered_fp <- combined_fp_injury %>% 
  select(Athlete, `Test Date`, incident_date, incident_sport, DaysDifference, issue_description, body_area, issue_description,
         area_of_body, `Jump Height (Flight Time) [cm]`, `Flight Time [ms]`, `RSI-modified [m/s]`, 
         `Flight Time:Contraction Time`, `Eccentric Peak Force Asymmetry [% L,R]`, 
         `Eccentric Deceleration RFD Asymmetry [% L,R]`, `Eccentric Deceleration RFD [N/s]`, 
         `Eccentric Acceleration Phase Duration [s]`, `Braking Phase Duration [s]`, `Peak Power / BM [W/kg]`, 
         `Concentric Peak Force / BM [N/kg]`, `Concentric Mean Force / BM [N/kg]`, `Concentric RFD Asymmetry [% L,R]`,
         `Eccentric Peak Force / BM [N/kg]`, `Eccentric Mean Power / BM [W/kg]`)

# Create Baseline
# Compare Month prior to Baselines

# Separating Data frames into Baseline and Comparisons
baseline <- filtered_fp %>% 
  filter(DaysDifference > 30)

prior_month <- filtered_fp %>% 
  filter(DaysDifference <= 30)

# Adds Identifier to Force Plate Metric Columns
baseline <- baseline %>% 
  rename_at(vars(9:ncol(baseline)), ~paste("var_", ., sep = ""))

# Selects median observation for each incident by athlete
median_per_injury <- baseline %>%
  group_by(Athlete, incident_date) %>%
  summarise(across(starts_with("var_"), list(median = ~median(.))))

# Remove Identifiers
baseline <- baseline %>% 
  rename_at(vars(starts_with("var_")), ~str_remove(., "var_"))

median_per_injury <- median_per_injury %>% 
  rename_at(vars(starts_with("var_")), ~str_remove(., "var_"))

median_per_injury <- median_per_injury %>% 
  rename_at(vars(ends_with("_median")), ~str_remove(., "_median"))


# Identifier for Median
median_per_injury <- median_per_injury %>% 
  rename_at(vars(3:ncol(median_per_injury)), ~paste("med_", ., sep = ""))

# Join the data frames
median_and_prior_df <- merge(prior_month, median_per_injury, by = c("Athlete", "incident_date"), all = TRUE)

# Remove Observations without Prior Month Jumps
median_and_prior_df <- median_and_prior_df %>% 
  filter(!is.na(`Test Date`))

# Remove Obsevations without Medians
median_and_prior_df <- median_and_prior_df %>% 
  filter(!is.na(`med_Jump Height (Flight Time) [cm]`))


# Finds % decrease value of Median versus the Prior Month Jump
median_and_prior_df <- median_and_prior_df %>%
  mutate(across(starts_with("med_"), 
                list(pct_decrease = ~ 100 * (.- get(sub("med_", "", cur_column()))) / .),
                .names = "pct_{.col}"))

# Format it correctly
median_and_prior_df <- median_and_prior_df %>%
  mutate(across(starts_with("pct_"), ~ . * -1))

# Final Dataframe with All Sports, FP data, and Injuries
# median_and_prior_df
```

## R Shiny Application with Toggle for Sport and Metric

```{r}
# Add Mechanism

ui <- fluidPage(
  selectInput("metric", "Choose Y-axis Metric", choices = sort(colnames(median_and_prior_df)[39:53])),
  selectInput("incident_sport", "Choose Sport", choices = c("Overall", sort(unique(median_and_prior_df$incident_sport)))),
  selectInput("body_area", "Choose Injury", choices = c("Overall", sort(unique(median_and_prior_df$body_area)))),
  selectInput("area_of_body", "Choose Area of Body", choices = c("Overall", sort(unique(median_and_prior_df$area_of_body)))),
  plotOutput("scatterPlot")
)

server <- function(input, output, session) {
  
  output$scatterPlot <- renderPlot({
    
  filtered_data <- median_and_prior_df %>%
    filter(
      (input$incident_sport == "Overall" | incident_sport == input$incident_sport) &
      (input$body_area == "Overall" | body_area == input$body_area) &
      (input$area_of_body == "Overall" | area_of_body == input$area_of_body))
      
    ggplot(data = filtered_data, aes_string(x = "DaysDifference", y = paste0("`", input$metric, "`"))) +
      geom_line(y = 0, col = "red", size = .5, alpha = .25) +
      geom_point() +
      geom_smooth(method = "lm", se = TRUE, color = "blue") +
      labs(title = paste("Correlation with Days before Injury and", input$metric),
           subtitle = paste("Sport:", input$incident_sport, "   ", 
                            "Injury:", input$body_area,  "   ", 
                            "Body Area:", input$area_of_body)) +
      theme_minimal() +
      scale_x_reverse()
  })
  
  # Fixes Closing Issues
    session$onSessionEnded(function() {
    stopApp()
  })

    
}



shinyApp(ui, server)

# grep("dif_", colnames(median_and_prior_df), value = TRUE)

# Flight Time and Jump Height for Basketball look good
# Peak Power for LAX/Soccer negative before, and increases towards injury
```

## Basketball Relationship

```{r}

bball_median_df <- median_and_prior_df %>% 
  filter(incident_sport == "Mens Basketball" | incident_sport == "Womens Basketball")

# dif_med_Jump Height (Flight Time) [cm]
# dif_med_Flight Time [ms]


# Chart/Regression
# ggplot(data = bball_median_df, aes(x = DaysDifference, y = `pct_med_Jump Height (Flight Time) [cm]`)) +
#   geom_point() +
#   geom_smooth(method = "lm", se = TRUE, color = "blue") +
#   scale_x_reverse()

# bball_model <- lm(`pct_med_Jump Height (Flight Time) [cm]` ~ DaysDifference, data = bball_median_df)
# summary(bball_model)
# confint(bball_model, level = .95)



# Select Relevant Columns
bball_median_df <- bball_median_df %>% 
  select(colnames(bball_median_df)[1:8], `pct_med_Jump Height (Flight Time) [cm]`, `pct_med_Flight Time [ms]`)

# Separate into another data frame to see if Three Day Grouping is better or not
three_day_bball <- bball_median_df

# Another dataframe for just Ankle Injuries
ankle_bball <- bball_median_df %>% 
  filter(body_area == "Ankle")

# Separate ankle three day average
three_day_ankle <- ankle_bball

# Get average by days difference
bball_median_df <- bball_median_df %>% 
  group_by(DaysDifference) %>% 
  mutate(averaged_Flight_Time = mean(`pct_med_Flight Time [ms]`)) %>% 
  mutate(averaged_Jump_Height = mean(`pct_med_Jump Height (Flight Time) [cm]`)) %>% 
  ungroup()

# Get average by days for ankles
ankle_bball <- ankle_bball %>% 
  group_by(DaysDifference) %>% 
  mutate(averaged_Flight_Time = mean(`pct_med_Flight Time [ms]`)) %>% 
  mutate(averaged_Jump_Height = mean(`pct_med_Jump Height (Flight Time) [cm]`)) %>% 
  ungroup()

# Get average by 3 day groupings
three_day_bball <- three_day_bball %>% 
  mutate(grouped_time = (DaysDifference - 1) %/% 3) %>% 
  mutate(grouped_time = ifelse(grouped_time == -1, 0, grouped_time)) %>% 
  group_by(grouped_time) %>% 
  mutate(averaged_Flight_Time = mean(`pct_med_Flight Time [ms]`)) %>% 
  mutate(averaged_Jump_Height = mean(`pct_med_Jump Height (Flight Time) [cm]`)) %>% 
  ungroup()

# Get average by 3 day groupings
three_day_ankle <- three_day_ankle %>% 
  mutate(grouped_time = (DaysDifference - 1) %/% 3) %>% 
  mutate(grouped_time = ifelse(grouped_time == -1, 0, grouped_time)) %>% 
  group_by(grouped_time) %>% 
  mutate(averaged_Flight_Time = mean(`pct_med_Flight Time [ms]`)) %>% 
  mutate(averaged_Jump_Height = mean(`pct_med_Jump Height (Flight Time) [cm]`)) %>% 
  ungroup()


# All Periods Charting
p1 <- ggplot(data = bball_median_df, aes(x = DaysDifference, y = averaged_Jump_Height)) +
  geom_line() +
  geom_line(y = 0, color = "blue") +
  geom_line(y = mean(bball_median_df$averaged_Jump_Height), color = "red", linetype = "dotted") + 
  scale_x_reverse() +
  theme_minimal() +
  labs(x = "",
       y = "% Change in Jump Height",
       title = "Jump Height Decreases Prior to Injury",
       subtitle = "(For Mens and Womens Basketball)") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, vjust = 0.5),
        text = element_text(size = 20)) +
  annotate("text", x = 1, y = -1.75, 
           label = paste("Mean:", round(mean(bball_median_df$averaged_Jump_Height), 2)),
           fontface = "bold", size = 5)

# Ankle Charting
p2 <- ggplot(data = ankle_bball, aes(x = DaysDifference, y = averaged_Jump_Height)) +
  geom_line() +
  geom_line(y = 0, color = "blue") +
  geom_line(y = mean(ankle_bball$averaged_Jump_Height), color = "red", linetype = "dotted") + 
  scale_x_reverse() +
  theme_minimal() +
  labs(x = "",
       y = "",
       title = "Jump Height Decreases Prior to Ankle Injury",
       subtitle = "(For Mens and Womens Basketball)") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, vjust = 0.5),
        text = element_text(size = 20)) +
  annotate("text", x = 1.25, y = -2, 
           label = paste("Mean:", round(mean(ankle_bball$averaged_Jump_Height), 2)),
           fontface = "bold", size = 5)

# 3 Day Periods
p3 <- ggplot(data = three_day_bball, aes(x = DaysDifference, y = averaged_Jump_Height)) +
  geom_line() +
  geom_line(y = 0, color = "blue") +
  geom_line(y = mean(three_day_bball$averaged_Jump_Height), color = "red", linetype = "dotted") +
  scale_x_reverse() +
  theme_minimal() +
  labs(x = "Days Until Injury",
       y = "% Change in Jump Height",
       title = "Jump Height Decreases Prior to Injury",
       subtitle = "(For Mens and Womens Basketball)") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, vjust = 0.5),
        text = element_text(size = 20)) +
  annotate("text", x = 1, y = -1.35, 
           label = paste("Mean:", round(mean(three_day_bball$averaged_Jump_Height), 2)),
           fontface = "bold", size = 5)

# 3 Day Periods for Ankle
p4 <- ggplot(data = three_day_ankle, aes(x = DaysDifference, y = averaged_Jump_Height)) +
  geom_line() +
  geom_line(y = 0, color = "blue") +
  geom_line(y = mean(three_day_ankle$averaged_Jump_Height), color = "red", linetype = "dotted") +
  scale_x_reverse() +
  theme_minimal() +
  labs(x = "Days Until Injury",
       y = "",
       title = "Jump Height Decreases Prior to Ankle Injury",
       subtitle = "(For Mens and Womens Basketball)") +
  theme(plot.title = element_text(hjust = 0.5, vjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, vjust = 0.5),
        text = element_text(size = 20)) +
  annotate("text", x = 1.25, y = -2.15, 
           label = paste("Mean:", round(mean(three_day_ankle$averaged_Jump_Height), 2)),
           fontface = "bold", size = 5)

# Run all of these at the same time
grid.arrange(p1, p2, p3, p4, ncol = 2)
grid.lines(x = c(0.51, 0.51, 1.51, 1.51), y = c(0, 2), gp = gpar(col = "black"))
grid.lines(x = c(0, 2), y = c(0.52, 0.52, 1.52, 1.52), gp = gpar(col = "black"))

```

## Proportional Hazards Model

```{r}

# Select Relevant Variables
df_coxph <- combined_fp_injury %>% 
  select(incident_date, `Test Date`, athlete_name, DaysDifference, ClosestTestToInjury, 
         incident_sport, body_area, `RSI-modified [m/s]`, `Flight Time [ms]`, `Jump Height (Flight Time) [cm]`)


# Separating Data into different timeframes
df_coxph <- df_coxph %>% 
  mutate(split_timeframe = ifelse(DaysDifference <= 90, 1, 0))

df_coxph <- df_coxph %>% 
  mutate(averaged_difference = ifelse(split_timeframe == 0, DaysDifference - 90, DaysDifference))

# Setting incident sport as Factor
df_coxph$incident_sport <- factor(df_coxph$incident_sport, levels = unique(df_coxph$incident_sport))

# Setting Reference Level
df_coxph$incident_sport <- relevel(df_coxph$incident_sport, ref = "Mens Basketball")

# Fit Cox Proportional Hazards model
cox_model <- coxph(Surv(averaged_difference, split_timeframe) ~ incident_sport + 
                     `Flight Time [ms]` + `Jump Height (Flight Time) [cm]`, data = df_coxph)

summary(cox_model)


# Assuming cox_model is the result of coxph()
survival_curves <- survfit(cox_model)
plot(survival_curves, main = "Survival Curves")






# Try it Basketball Specific


bb_coxph <- df_coxph %>% 
  filter(incident_sport == "Mens Basketball" | incident_sport == "Womens Basketball")

# Setting incident sport as Factor
bb_coxph$incident_sport <- factor(bb_coxph$incident_sport, levels = unique(bb_coxph$incident_sport))

# Setting Reference Level
bb_coxph$incident_sport <- relevel(bb_coxph$incident_sport, ref = "Mens Basketball")



# Fit Cox Proportional Hazards model
bb_cox_model <- coxph(Surv(averaged_difference, split_timeframe) ~ incident_sport + 
                     `Flight Time [ms]` + `Jump Height (Flight Time) [cm]`, data = bb_coxph)

summary(bb_cox_model)


# Assuming bb_cox_model is the result of coxph()
survival_curves <- survfit(bb_cox_model)
plot(survival_curves, main = "Survival Curves")

```
