---
title: "Track & XC"
author: "Parker Booth"
date: "2024-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
```{r}
library(dplyr) # Data Manipulation Package %>% 
library(tidyr) # Data Cleaning Package
library(readxl) # Excel
library(tidyverse) # Cleaning/Manipulating
library(purrr) # Initializing Lists to Data Frames
library(lubridate) # Working with Dates
library(openxlsx) # Excel 
library(stringr) # String Splitting
library(ggplot2) # Plotting
library(stringr) # Prefix/Suffix Editing
library(manipulate) # Easy Chart Manipulation
library(shiny) # Interactive Charting
library(gridExtra) # Plotting Charts Together
library(grid) # Grid Manipulation
library(ggcorrplot) # Correlation Chart
library(cowplot) # Combining Charts
library(lme4) # Hierarchical Modeling
library(lmerTest) # More Hierarchical Modeling
library(fmsb) # Radar Charts
```

## Initial Data Cleaning
```{r}
# Set Working Directory 
setwd("Path")

# Read in Force Plate Data
xc_fp <- read_excel("Path")
dj_top <- read_excel("Path")
dj <- read_excel("Path", skip = 7)
track_fp <- read_excel("Path", skip = 7)


# Setting Date Formats
xc_fp$`Test Date` <- as.Date(xc_fp$`Test Date`, format = "%b %d, %Y")
dj_top$`Test Date` <- as.Date(dj_top$`Test Date`, format = "%b %d, %Y")
dj$`Test Date` <- as.Date(dj$`Test Date`, format = "%b %d, %Y")
track_fp$`Test Date` <- as.Date(track_fp$`Test Date`, format = "%b %d, %Y")


# # Code to select N/A values when Needed
# # Step 1: Count the number of NA values in each column
# na_counts <- sapply(xc_fp, function(x) sum(is.na(x)))
# 
# # Convert na_counts to a data frame for easier manipulation with dplyr
# na_counts_df <- data.frame(column = names(na_counts), na_count = na_counts)
# 
# # Step 2: Filter columns with NA values > 0
# columns_with_na <- na_counts_df %>%
#   filter(na_count > 0)
# 
# # OR DO THIS
# na_columns <- xc_fp %>%
#   select(where(~ any(is.na(.))))

# Read in Rosters
tk_roster <- read_excel("Path")
xc_roster <- read_excel("Path")

# Combine into one roster
roster <- rbind(tk_roster, xc_roster)

# Add Roster Data to all Data Frames
xc_fp <- merge(xc_fp, roster, by = "Athlete", all.x = TRUE)

xc_fp <- xc_fp %>% 
  mutate(Distance_Category = case_when(
    Event %in% c("LD", "5K/15") ~ "Long_Distance",
    Event %in% c("MD", "15/5K") ~ "Medium_Distance",
    TRUE ~ "Unknown"
  ))

xc_fp <- xc_fp %>% 
  select(Athlete, `Test Type`, `Test Date`, Trial, Event, Distance_Category, Status, Team, everything())

dj_top <- merge(dj_top, roster, by = "Athlete", all.x = TRUE)
dj_top <- dj_top %>% 
  select(Athlete, `Test Type`, `Test Date`, `Key Trial`, Event, Status, Team, everything())

dj <- merge(dj, roster, by = "Athlete", all.x = TRUE)
dj <- dj %>% 
  select(Athlete, `Test Type`, `Test Date`, Trial, Event, Status, Team, everything())

track_fp<- merge(track_fp, roster, by = "Athlete", all.x = TRUE)

track_fp <- track_fp %>% 
  mutate(Distance_Category = case_when(
    Event == "400M/800M" ~ "Medium_Distance",
    Event %in% c("Hurdles", "Sprints") ~ "Short_Distance",
    Event == "Jumps" ~ "Jumps",
    Event == "Multis" ~ "Multis",
    TRUE ~ "Unknown"
  ))

track_fp <- track_fp %>% 
  select(Athlete, `Test Type`, `Test Date`, Trial, Event, Distance_Category, Status, Team, everything())





# See differences between XC & Track colnames
track_col <- colnames(track_fp)
xc_col <- colnames(xc_fp)

unique_in_vec1 <- setdiff(track_col, xc_col)
unique_in_vec2 <- setdiff(xc_col, track_col)

# Rename a column to make it match
names(track_fp)[names(track_fp) == "Concentric Impulse (Abs) / BM [N s]"] <- "Concentric Impulse (Abs) / BM [N s/kg]"

# Combine XC & Track fp data frames
fp <- rbind(xc_fp, track_fp)


# Writing Cleaned Files
setwd("C:/Users/spike/Box/AHPS Projects/AHPS Track XC Project/Exported Cleaned Files")

write.xlsx(fp, "Combined CMJ Force Plate Data.xlsx", rowNames = FALSE)
write.xlsx(roster, "Combined Roster.xlsx", rowNames = FALSE)
write.xlsx(xc_fp, "XC CMJ Force Plate Data.xlsx", rowNames = FALSE)
write.xlsx(track_fp, "Track CMJ Force Plate Data.xlsx", rowNames = FALSE)
write.xlsx(dj, "DJ Force Plate Data.xlsx", rowNames = FALSE)
write.xlsx(dj_top, "Top Trial DJ Force Plate Data.xlsx", rowNames = FALSE)
```

## Initial Exploratory Analysis
```{r}

# Initial Plotting
ggplot(fp, aes(x = `Jump Height (Flight Time) [cm]`, y = `Peak Power / BM [W/kg]`, color = Team)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Relationship between Jump Height & Peak Power (Using all force plate data)") +
  facet_wrap(~ Distance_Category)

# Re-order to easily see RSI Modified
fp <- fp %>% 
 select(Athlete, `Test Type`, `Test Date`, Trial, Event, Distance_Category, Status, Team, `RSI-modified [m/s]`, everything())

# Create a data frame based on the best RSI-Mod in the trials 
best_fp <- fp %>% 
  group_by(`Test Date`, Athlete) %>%
  slice_max(`RSI-modified [m/s]`, n = 1, with_ties = FALSE) %>%
  ungroup()

# # Writing Cleaned Files
# setwd("C:/Users/spike/Box/AHPS Projects/AHPS Track XC Project/Exported Cleaned Files")
# 
# write.xlsx(best_fp, "Combined CMJ Force Plate Data (Best Jumps).xlsx", rowNames = FALSE)


# Plotting with best trial taken
ggplot(best_fp, aes(x = `Jump Height (Flight Time) [cm]`, y = `Peak Power / BM [W/kg]`, color = Team)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Relationship between Jump Height & Peak Power (Using best jumps)") +
  facet_wrap(~ Distance_Category)

# More Plotting
ggplot(best_fp, aes(x = `RSI-modified [m/s]`, y = `Concentric Peak Force / BM [N/kg]`, color = Team)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Relationship between RSI Mod & Concentric Peak Force (Using best jumps)") +
  facet_wrap(~ Distance_Category)

ggplot(best_fp, aes(x = `RSI-modified [m/s]`, y = `Eccentric Peak Force / BM [N/kg]`, color = Team)) +
  geom_point(alpha = 0.1) +
  geom_smooth(method = "lm", se = FALSE) +
  theme_minimal() +
  labs(title = "Relationship between RSI Mod & Eccentric Peak Force (Using best jumps)") +
  facet_wrap(~ Distance_Category)


```

## Track & XC Middle Distance Comparison using RShiny
```{r}

# Select just the medium distance group in a data frame (Only Group with XC & Track)
med_best <- best_fp %>% 
  filter(Distance_Category == "Medium_Distance" | Athlete %in% c("Athlete 1", "Athlete 2", "Athlete 3"))




# Define UI for the app
ui <- fluidPage(
  titlePanel("Interactive ggplot with Shiny"),
  sidebarLayout(
    sidebarPanel(
      selectInput("yvar", "Y-axis variable:", choices = names(med_best)),
      tableOutput("selectedPoints")
    ),
    mainPanel(
      plotOutput("plot", brush = brushOpts(id = "plot_brush"))
    )
  )
)


server <- function(input, output) {
  # Reactive expression to create the ggplot
  plotData <- reactive({
    y <- sym(input$yvar)
    ggplot(med_best, aes(x = Team, y = !!y, color = Team)) +
      geom_boxplot() +
      theme_minimal()
  })
  
  # Render the plot
  output$plot <- renderPlot({
    plotData()
  })
  

}


# Run the app
shinyApp(ui = ui, server = server)


```

## Run Heirarchical Linear Model on Interesting Variables
```{r}
# Select just the medium distance group in a data frame (Only Group with XC & Track)
med_best <- best_fp %>% 
  filter(Distance_Category == "Medium_Distance" | Athlete %in% c("Athlete 1", "Athlete 2", "Athlete 3"))



# Checking to see if variation exists between the groups
rsi_initial_model <- lmer(`RSI-modified [m/s]` ~ 1 + (1 | Team), data = med_best)
summary(rsi_initial_model)

eccentric_model <- lmer(`Eccentric Mean Power / BM [W/kg]` ~ 1 + (1 | Team), data = med_best)
summary(eccentric_model)

concentric_model <- lmer(`Concentric Mean Power / BM [W/kg]` ~ 1 + (1 | Team), data = med_best)
summary(concentric_model)


ranova(rsi_initial_model)
ranova(eccentric_model)
ranova(concentric_model)








# Define the function to calculate ICC
calculate_icc <- function(var) {
  # Fit the intercept-only model
  model <- lmer(var ~ (1 | Team), data = med_best)
  
  # Extract variance components
  var_components <- as.data.frame(VarCorr(model))
  
  # Calculate ICC
  icc <- var_components$vcov[1] / (var_components$vcov[1] + var_components$vcov[2])
  
  return(icc)
}

# Get the column names of your data set
example_df <- med_best %>% 
  select_if(~!all(is.na(.)))
  
all_column_names <- colnames(example_df)

# Drop the specific column name from the variables vector
variables <- all_column_names[9:length(all_column_names)]


# Initialize an empty data frame to store the results
icc_results <- data.frame(Variable = character(), ICC = numeric(), stringsAsFactors = FALSE)

# Loop through each variable
for (variable in variables) {
  print(variable)
  icc <- calculate_icc(med_best[[variable]])
  icc_results <- rbind(icc_results, data.frame(Variable = variable, ICC = icc))
}

icc_results <- icc_results %>%
  mutate(ICC = format(ICC, scientific = FALSE)) %>% 
  mutate(ICC = round(as.numeric(ICC), 2))


icc_results <- icc_results %>%
  filter(!str_detect(Variable, regex("landing", ignore_case = TRUE))) %>% 
  filter(!str_detect(Variable, regex("Right", ignore_case = TRUE))) %>% 
  filter(!str_detect(Variable, regex("Left", ignore_case = TRUE)))

```

## Comparing metrics between whole team
```{r}
best_xc_fp <- best_fp %>% 
  filter(Team == "XC")

best_track_fp <- best_fp %>% 
  filter(Team == "TF")

ggplot(best_xc_fp, aes(x = Distance_Category, y = `RSI-modified [m/s]`, color = Distance_Category)) +
  geom_boxplot() +
  theme_minimal()

ggplot(best_track_fp, aes(x = Distance_Category, y = `RSI-modified [m/s]`, color = Distance_Category)) +
  geom_boxplot() +
  theme_minimal()




bestish_fp <- best_fp %>% 
  mutate(Teamish = paste0(Team, Distance_Category))

# Define the specific order you want for the column
specific_order <- c("TFJumps", "TFMultis", "TFShort_Distance", "TFMedium_Distance", "XCMedium_Distance", "XCLong_Distance", "XCUnknown")

# Convert the column to a factor with the specified order
bestish_fp <- bestish_fp %>%
  mutate(Teamish = factor(Teamish, levels = specific_order))

ggplot(bestish_fp, aes(x = Teamish, y = `RSI-modified [m/s]`, color = Team)) +
  geom_boxplot() +
  theme_minimal()



ordering_here <- c("Order Of Athletes by Team")

best2_fp <- best_fp %>% 
  mutate(Athlete = factor(Athlete, levels = ordering_here))


ggplot(best2_fp, aes(x = Athlete, y = `RSI-modified [m/s]`, color = Team)) +
  geom_boxplot() +
  labs(
    title = "",
    x = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

med2_best <- med_best %>% 
  mutate(Athlete = factor(Athlete, levels = ordering_here))

ggplot(med2_best, aes(x = Athlete, y = `RSI-modified [m/s]`, color = Team)) +
  geom_boxplot() +
  labs(
    title = "",
    x = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



column_names <- colnames(best_fp)
concentric_columns <- column_names[grepl("Concentric", column_names)]

# Print the matching column names
print(concentric_columns)

ggplot(med2_best, aes(x = Athlete, y = `Concentric RFD / BM [N/s/kg]`, color = Team)) +
  geom_boxplot() +
  labs(
    title = "",
    x = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



# Trying a Radar Charting and lollipop charting between teams on some metrics


# Selecting the 5 variables I want to investigate
med2_best <- med2_best %>% 
  select(Athlete, `Test Date`, Team, `RSI-modified [m/s]`, `Eccentric Mean Force [N]`, 
         `Concentric Mean Force [N]`, `Peak Power [W]`, `Flight Time [ms]`)

# Try Team average charts
cols_with_brackets <- colnames(med2_best)[grepl("\\[", colnames(med2_best))]
team_avg <- med2_best %>% 
  group_by(Team) %>%
  summarize(across(all_of(cols_with_brackets), mean, na.rm = TRUE))

team_avg <- team_avg %>% 
  column_to_rownames("Team")

team_avg <- rbind(
  max = c(0.6, 600, 1200, 3000, 600),  # Example maximum values for each variable
  min = c(0, 500, 1000, 2500, 500),    # Example minimum values for each variable
  team_avg
)


# RUN THIS ALL AT THE SAME TIME OR IT WILL NOT WORK

colors_border <- c("#FF6347", "#1E90FF")  # Colors for TF and XC
colors_in <- c("#FF634780", "#1E90FF80")  # Transparent colors for filling


# Create radar chart
radarchart(team_avg,
           axistype = 1,        # 1: Different scales for each variable
           pcol = colors_border, # Specify colors for each group
           pfcol = colors_in,
           plwd = 2,            # Line width
           plty = 3,            # Line type
           axislabcol = "black",# Color of axis labels
           vlcex = 0.7,         # Label cex
           cglcol = "grey",     # Color of grid lines
           cglty = 1,           # Grid line type
           cglwd = 0.8,         # Grid line width
           calcex = 0.4,
           title = "Comparison of TF and XC Groups"  # Chart title
)

# Add a legend
legend(x=0.9, y=1.4, legend = rownames(team_avg[-c(1,2),]), bty = "n", pch=20 , col=colors_in , text.col = "grey", cex=1.2, pt.cex=3)


custom_labels <- c("Jumps", "Multis", "Short Dist", "Medium Dist", "Medium Dist", "Long Dist", "Unknown")


ggplot(bestish_fp, aes(x = Teamish, y = `RSI-modified [m/s]`, color = Team)) +
  geom_boxplot() +
    scale_x_discrete(labels = custom_labels) + # Set custom x-axis labels
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text( size = 12),
    axis.title.x = element_blank(),
    legend.text = element_text(size = 14),       # Make legend text larger
    legend.title = element_text(size = 16),
    axis.title.y = element_text(face = "bold", size = 14)
  )


```

## Adding in Running Times
```{r}
# Set your working directory to the folder containing your CSV files
setwd("Path")

# List all CSV files in the directory
csv_files <- list.files(pattern = "*.csv")

# Function to read a CSV file and standardize column types
read_and_standardize <- function(file) {
  df <- read.csv(file, stringsAsFactors = FALSE)
  df <- mutate_all(df, as.character) # Convert all columns to character type
  return(df)
}

# Read each CSV file with the standardization function
list_of_dfs <- lapply(csv_files, read_and_standardize)

# Combine all data frames into one
med_race_times <- bind_rows(list_of_dfs)
med_race_times <- distinct(med_race_times)

# Assuming med_race_times is your data frame
med_race_times <- med_race_times %>% 
  select(-X, -X.1) %>%  # Remove columns named X and X.1 if they are present
  mutate(Event = ifelse(Event == "600y", "600", Event))

med_race_times$Date <- as.Date(med_race_times$Date, format = "%d-%b-%y")


# Function to format times
format_times <- function(time_str) {
  if (str_detect(time_str, ":")) {
    # If time is already in MM:SS.ss format, return as is
    return(time_str)
  } else {
    # If time is in SS.ss format, convert to MM:SS.ss
    seconds <- as.numeric(time_str)
    minutes <- floor(seconds / 60)
    remaining_seconds <- seconds %% 60
    return(sprintf("%02d:%05.2f", minutes, remaining_seconds))
  }
}

med_race_times <- med_race_times %>%
  mutate(formatted_time = sapply(Time, format_times)) %>% 
  select(-Time) %>% 
  rename(Time = formatted_time)


convert_to_numeric <- function(time_str) {
  if (str_detect(time_str, ":")) {
    parts <- str_split(time_str, ":", simplify = TRUE)
    return(as.numeric(parts[1]) * 60 + as.numeric(parts[2]))
  } else {
    return(as.numeric(time_str))
  }
}

med_race_times <- med_race_times %>%
  mutate(NumericTime = sapply(Time, convert_to_numeric))

med_race_times <- med_race_times %>% 
  mutate(EventNumeric = case_when(
    Event == "Mile" ~ "1600",
    Event == "SMR" ~ "1600",
    Event == "DMR" ~ "4000",
    Event == "4x400" ~ "1600",
    Event == "4x800" ~ "3200",
    Event == "5k" ~ "5000",
    Event == "6k" ~ "6000",
    Event == "4x200" ~ "800",
    Event == "4x100" ~ "400",
    Event == "3000S" ~ "3000",
    Event == "4x1600" ~ "6400",
    Event == "4k" ~ "4000",
    TRUE ~ as.character(Event)
  )) %>% 
  mutate(EventNumeric = as.numeric(EventNumeric))

med_race_times$relay <- ifelse(grepl("x|R", med_race_times$Event), "relay", "no")

# Optionally, you can write the combined data frame to a new CSV file
setwd("Path")
write.csv(med_race_times, "med_race_times.csv", row.names = FALSE)

```

## Start Comparing Running Times with Force Plates (Built the combined data frame here)
```{r}
med_best <- best_fp %>% 
  filter(Distance_Category == "Medium_Distance" | Athlete %in% c("Athlete 1", "Athlete 2", "Athlete 3"))

true_med_times <- med_race_times %>% 
  filter(EventNumeric >= 400 & EventNumeric <= 3000)

# Find the name with the fastest time for each event
fastest_times <- true_med_times %>%
  group_by(Event) %>%
  slice(which.min(NumericTime)) %>%
  select(Athlete, Event, Time, relay)


# Plotting to show how close force plate tests are to races

# Test Data frames
med_fp_test <- med_best
med_times_test <- true_med_times

med_times_test$source <- "df1"
med_fp_test$source <- "df2"

med_fp_test <- med_fp_test %>% 
  select(Athlete, `Test Date`, source) %>% 
  rename(Date = `Test Date`)

med_times_test <- med_times_test %>% 
  select(Athlete, Date, source)

# Combine both data frames
combined_test_df <- rbind(med_fp_test, med_times_test)

# Plot using ggplot2
ggplot(combined_test_df, aes(x = Date, y = Athlete, color = source)) +
  geom_point(size = 3, alpha = .3) +
  scale_color_manual(values = c("df1" = "green", "df2" = "red"), labels = c("df1" = "Race", "df2" = "Test")) +
  labs(title = "Explanatory Plot of Dates and Names",
       x = "Date",
       y = "Name",
       color = "Data Source") +
  theme_minimal() +
  theme(legend.text = )



# Setting Nearest Race variable


# Filter out dates before 2021
med_best <- med_best %>%
  filter(`Test Date` >= as.Date("2021-01-01")) %>% 
  rename(Date = `Test Date`)

true_med_times <- true_med_times %>%
  filter(Date >= as.Date("2021-01-01"))


true_med_times <- true_med_times %>%
  rename(Race.Date = Date)

med_best <- med_best %>% 
  rename(Test.Date = Date)



# Function to get the closest test date prior to each race date
get_closest_test_date <- function(race_date, test_dates) {
  prior_test_dates <- test_dates[test_dates < race_date]
  if (length(prior_test_dates) == 0) {
    return(NA)
  }
  return(max(prior_test_dates))
}

# List of unique athletes
athletes <- unique(true_med_times$Athlete)

# Initialize an empty data frame to store results
result <- data.frame()

# Loop through each athlete
for (athlete in athletes) {
  # Filter data for the current athlete
  race_df <- true_med_times %>% filter(Athlete == athlete)
  test_df <- med_best %>% filter(Athlete == athlete)
  
  # Get the closest test date for each race date
  race_df$ClosestTestDate <- sapply(race_df$Race.Date, get_closest_test_date, test_dates = test_df$Test.Date)
  
  # Append the result
  result <- bind_rows(result, race_df)
}

# Print the final result
print(result)

result$ClosestTestDate <- as.Date(result$ClosestTestDate, format = "%b %d, %Y")
result$DaysBeforeRace <- as.numeric(difftime(result$Race.Date, result$ClosestTestDate, units = "days"))

result <- result %>% 
  filter(!is.na(ClosestTestDate)) %>% 
  mutate(Reliability = case_when(
    DaysBeforeRace <= 3 ~ "Amazing",
    DaysBeforeRace <= 10 ~ "Good",
    DaysBeforeRace <= 25 ~ "Poor",
    TRUE ~ "Bad")) %>% 
  rename(Test.Date = ClosestTestDate)



# Merge into one Data frame
race_fp_med <- left_join(result, med_best, by = c("Athlete", "Test.Date"))

write.xlsx(race_fp_med, "Race & FP combined.xlsx", rowNames = FALSE)

```

## Analysis with Running and Force Plate
```{r}
race_fp_med <- race_fp_med %>%
  filter(relay != "relay") %>% 
  filter(`RSI-modified [m/s]` <= .9)

# Define UI for the app
ui <- fluidPage(
  titlePanel("Interactive ggplot with Shiny"),
  sidebarLayout(
    sidebarPanel(
      selectInput("yvar", "Y-axis variable:", choices = names(race_fp_med)),
      tableOutput("selectedPoints")
    ),
    mainPanel(
      plotOutput("plot", brush = brushOpts(id = "plot_brush"))
    )
  )
)

server <- function(input, output) {
  # Reactive expression to create the ggplot
  plotData <- reactive({
    y <- sym(input$yvar)
    ggplot(race_fp_med, aes(x = NumericTime, y = !!y, color = Team)) +
      geom_point() +
      geom_smooth(method = "lm", se = FALSE) +  # Adding a trend line with linear model
      facet_wrap(~ Reliability) +
      theme_minimal()
  })
  
  # Render the plot
  output$plot <- renderPlot({
    plotData()
  })
}

# Run the app
shinyApp(ui = ui, server = server)


# Try by fastest athlete
erickson_testing <- race_fp_med %>% 
  filter(Athlete == "Athlete 1")

Avg.Reliability <-  mean(erickson_testing$DaysBeforeRace)

ggplot(erickson_testing, aes(x = NumericTime, y = `RSI-modified [m/s]`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Adding a trend line with linear model
  theme_minimal() +
  annotate("text", x = Inf, y = Inf, label = paste("Avg # of Days (Race - Test): ", round(Avg.Reliability, 2)), 
         hjust = 1.2, vjust = 5.5, size = 5, color = "red") +
  facet_wrap(~Reliability)




# Initialize an empty data frame to store results
results <- data.frame(Athlete = character(), Slope = numeric(), AvgReliability = numeric(), stringsAsFactors = FALSE)

# Loop through each athlete
for (athlete in unique(race_fp_med$Athlete)) {
  
  athlete_data <- race_fp_med %>% 
    filter(Athlete == athlete)
  
  lm_model <- lm(`Peak Power / BM [W/kg]` ~ NumericTime, data = athlete_data)
  slope <- coef(lm_model)["NumericTime"]
  
  # Calculate average reliability
  avg_reliability <- round(mean(athlete_data$DaysBeforeRace, na.rm = TRUE), 2)
  
  # Store the results in the data frame
  results <- rbind(results, data.frame(Athlete = athlete, Slope = slope, AvgReliability = avg_reliability))
}

print(results)




# Try out some more Hierarchical Modeling

race_800 <- race_fp_med %>% 
  filter(EventNumeric == 800)
  
# Sort the data frame by the Time column in descending order
race_800 <- race_800 %>% 
  arrange(NumericTime)

race_800 <- race_800 %>% 
  mutate(Fastest_or_Not = ifelse(row_number() <= nrow(race_800) / 4, "High", "Low")) %>% 
  select(Fastest_or_Not, everything())

race_800 <- race_800 %>% 
  mutate(Slowest_25 = ifelse(row_number() >= (nrow(race_800) - nrow(race_800) / 4), "Slower", "Quick")) %>% 
  select(Slowest_25, everything())

# Initial Attempt at a model
hier_800_model <- lmer(`RSI-modified [m/s]` ~ 1 + (1 | Fastest_or_Not), data = race_800)
summary(concentric_model)




# Define the function to calculate ICC
calculate_icc <- function(var) {
  # Fit the intercept-only model
  model <- lmer(var ~ (1 | Fastest_or_Not), data = race_800)
  
  # Extract variance components
  var_components <- as.data.frame(VarCorr(model))
  
  # Calculate ICC
  icc <- var_components$vcov[1] / (var_components$vcov[1] + var_components$vcov[2])
  
  return(icc)
}

# Get the column names of your data set
example_df <- race_800 %>% 
  select_if(~!all(is.na(.)))
  
all_column_names <- colnames(example_df)

# Drop the specific column name from the variables vector
variables <- all_column_names[19:length(all_column_names)]


# Initialize an empty data frame to store the results
icc_results <- data.frame(Variable = character(), ICC = numeric(), stringsAsFactors = FALSE)

# Loop through each variable
for (variable in variables) {
  print(variable)
  icc <- calculate_icc(race_800[[variable]])
  icc_results <- rbind(icc_results, data.frame(Variable = variable, ICC = icc))
}

icc_results <- icc_results %>%
  mutate(ICC = format(ICC, scientific = FALSE)) %>% 
  mutate(ICC = round(as.numeric(ICC), 2))


icc_results <- icc_results %>%
  filter(!str_detect(Variable, regex("landing", ignore_case = TRUE))) %>% 
  filter(!str_detect(Variable, regex("Right", ignore_case = TRUE))) %>% 
  filter(!str_detect(Variable, regex("Left", ignore_case = TRUE)))

race_800 <- race_800 %>% 
  filter(`RSI-modified [m/s]` <= .8)

# Significant Decrease in the Peak Power once getting out of the top 25% of races
ggplot(data = race_800, aes(x = NumericTime, y = `Peak Power / BM [W/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(120), max(160), by = 2)) +
  theme_minimal() +
  labs(title = "800m Race") +
  theme(legend.position = "none")


ggplot(data = race_800, aes(x = NumericTime, y = `Force at Peak Power [N]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(120), max(160), by = 2)) +
  theme_minimal() +
  labs(title = "800m Race") +
  theme(legend.position = "none")

ggplot(data = race_800, aes(x = NumericTime, y = `RSI-modified [m/s]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(120), max(160), by = 2)) +
  theme_minimal() +
  labs(title = "800m Race") +
  theme(legend.position = "none")


# Do the same for the 1500 race
race_1500 <- race_fp_med %>% 
  filter(EventNumeric == 1500)
  
# Sort the data frame by the Time column in descending order
race_1500 <- race_1500 %>% 
  arrange(NumericTime)

race_1500 <- race_1500 %>% 
  mutate(Fastest_or_Not = ifelse(row_number() <= (nrow(race_1500) / 4), "High", "Low")) %>% 
  select(Fastest_or_Not, `Peak Power / BM [W/kg]`, everything())

ggplot(data = race_1500, aes(x = NumericTime, y = `RSI-modified [m/s]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(360), by = 5)) +
  theme_minimal() +
  labs(title = "1500m Race") +
  theme(legend.position = "none")

ggplot(data = race_1500, aes(x = NumericTime, y = `Eccentric Peak Force / BM [N/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(360), by = 5)) +
  theme_minimal() +
  labs(title = "1500m Race") +
  theme(legend.position = "none")

ggplot(data = race_1500, aes(x = NumericTime, y = `Peak Power / BM [W/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(360), by = 5)) +
  theme_minimal() +
  labs(title = "1500m Race") +
  theme(legend.position = "none")




ggplot(data = race_1600, aes(x = NumericTime, y = `RSI-modified [m/s]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(360), by = 5)) +
  theme_minimal() +
  labs(title = "1600m Race") +
  theme(legend.position = "none")

ggplot(data = race_1600, aes(x = NumericTime, y = `Eccentric Peak Force / BM [N/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(360), by = 5)) +
  theme_minimal() +
  labs(title = "1600m Race") +
  theme(legend.position = "none")

ggplot(data = race_1600, aes(x = NumericTime, y = `Peak Power / BM [W/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(360), by = 5)) +
  theme_minimal() +
  labs(title = "1600m Race") +
  theme(legend.position = "none")

ggplot(data = race_3000, aes(x = NumericTime, y = `RSI-modified [m/s]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(3600), by = 10)) +
  theme_minimal() +
  labs(title = "3000m Race") +
  theme(legend.position = "none")

ggplot(data = race_3000, aes(x = NumericTime, y = `Concentric Peak Force / BM [N/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(3600), by = 10)) +
  theme_minimal() +
  labs(title = "3000m Race") +
  theme(legend.position = "none")

ggplot(data = race_3000, aes(x = NumericTime, y = `Peak Power / BM [W/kg]`, color = Fastest_or_Not)) +
  geom_point() +
  scale_x_continuous(breaks = seq(min(200), max(3600), by = 10)) +
  theme_minimal() +
  labs(title = "3000m Race") +
  theme(legend.position = "none")

# Reliability Ratings
mean(race_800$DaysBeforeRace)
table(race_800$Reliability)


mean(race_1500$DaysBeforeRace)
table(race_1500$Reliability)

mean(race_1600$DaysBeforeRace)
table(race_1600$Reliability)


mean(race_3000$DaysBeforeRace)
table(race_3000$Reliability)




```

## Set up for Radar Charting
```{r}
# Working with race_800
# Take all tests from the top 2 athletes
radar800 <- race_800 %>% 
  mutate(top2 = ifelse(Athlete == "Athlete 1" | Athlete == "Athlete 2", "yes", "no")) %>% 
  mutate(bottom4 = ifelse(Athlete == "Athlete 3" | Athlete == "Athlete 4" | 
                          Athlete == "Athlete 5"| Athlete == "Athlete 6", "yes", "no"))

# Select relevant variables to look into
radar800 <- radar800 %>% 
  select(top2, bottom4, Athlete, Race.Date, NumericTime, EventNumeric, Test.Date, DaysBeforeRace, Reliability, Team, 
         `RSI-modified [m/s]`, `Peak Power / BM [W/kg]`, `Jump Height (Flight Time) [cm]`, 
         `Eccentric Peak Force / BM [N/kg]`, `Concentric Peak Force / BM [N/kg]`)


# Included top 2 racers only
radar800_best <- radar800 %>% 
  filter(top2 == "yes")

# Include Bottom Racers 
radar800worst <- radar800 %>% 
  filter(bottom4 == "yes")
  





radar800 <- radar800 %>%
  filter(Athlete != "Athlete 1")

# Weighted Average
wgt_avg_800 <- radar800 %>%
  group_by(Athlete) %>%
  summarise(
    wgt_RSI = weighted.mean(`RSI-modified [m/s]`, w = 1 / DaysBeforeRace),
    wgt_Peak_Power = weighted.mean(`Peak Power / BM [W/kg]`, w = 1 / DaysBeforeRace),
    wgt_Jump_Height = weighted.mean(`Jump Height (Flight Time) [cm]`, w = 1 / DaysBeforeRace),
    wgt_Ecc_Peak_Force = weighted.mean(`Eccentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace),
    wgt_Con_Peak_Force = weighted.mean(`Concentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace)
  )

wgt_team_avg <- radar800 %>% 
  group_by(Team) %>% 
  summarise(
    wgt_RSI = weighted.mean(`RSI-modified [m/s]`, w = 1 / DaysBeforeRace),
    wgt_Peak_Power = weighted.mean(`Peak Power / BM [W/kg]`, w = 1 / DaysBeforeRace),
    wgt_Jump_Height = weighted.mean(`Jump Height (Flight Time) [cm]`, w = 1 / DaysBeforeRace),
    wgt_Ecc_Peak_Force = weighted.mean(`Eccentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace),
    wgt_Con_Peak_Force = weighted.mean(`Concentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace)
  ) %>% 
  mutate(Athlete = ifelse(Team == "TF", "Track Average", "XC Average")) %>% 
  select(Athlete, Team, everything())

# Merge the top 2 for an average
wgt_avg_800_top2 <- radar800_best %>% 
  summarise(
    wgt_RSI = weighted.mean(`RSI-modified [m/s]`, w = 1 / DaysBeforeRace),
    wgt_Peak_Power = weighted.mean(`Peak Power / BM [W/kg]`, w = 1 / DaysBeforeRace),
    wgt_Jump_Height = weighted.mean(`Jump Height (Flight Time) [cm]`, w = 1 / DaysBeforeRace),
    wgt_Ecc_Peak_Force = weighted.mean(`Eccentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace),
    wgt_Con_Peak_Force = weighted.mean(`Concentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace)
  ) %>% 
  mutate(Athlete = "Best Average") %>% 
  mutate(Team = NA) %>% 
  select(Athlete, Team, everything())

# Merge bottom 4 for an average
wgt_avg_800_bottom4 <- radar800worst %>% 
  summarise(
    wgt_RSI = weighted.mean(`RSI-modified [m/s]`, w = 1 / DaysBeforeRace),
    wgt_Peak_Power = weighted.mean(`Peak Power / BM [W/kg]`, w = 1 / DaysBeforeRace),
    wgt_Jump_Height = weighted.mean(`Jump Height (Flight Time) [cm]`, w = 1 / DaysBeforeRace),
    wgt_Ecc_Peak_Force = weighted.mean(`Eccentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace),
    wgt_Con_Peak_Force = weighted.mean(`Concentric Peak Force / BM [N/kg]`, w = 1 / DaysBeforeRace)
  ) %>% 
  mutate(Athlete = "Lowest Average") %>% 
  mutate(Team = NA) %>% 
  select(Athlete, Team, everything())


# Combine Data Frames and re-introduce the teams
wgt_avg_800 <- merge(wgt_avg_800, roster)

wgt_avg_800 <- wgt_avg_800 %>% 
  select(-Event, -Status)

wgt_avg_800 <- rbind(wgt_team_avg, wgt_avg_800)
wgt_avg_800 <- rbind(wgt_avg_800_top2, wgt_avg_800)
wgt_avg_800 <- rbind(wgt_avg_800_bottom4, wgt_avg_800)


write.xlsx(wgt_avg_800, "Weighted FP Data for Radar Charts.xlsx")



```

## Radar Charting
```{r}
library(fmsb)

radarCharting <- wgt_avg_800

# Standardize numeric columns
numeric_cols <- c("wgt_RSI", "wgt_Peak_Power", "wgt_Jump_Height", "wgt_Ecc_Peak_Force", "wgt_Con_Peak_Force")

min_max_normalize <- function(x) {
  (x - min(x, na.rm = TRUE)) / (max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
}

# Apply min-max normalization to numeric columns
radarCharting[numeric_cols] <- lapply(radarCharting[numeric_cols], min_max_normalize)

averages <- radarCharting %>% 
  filter(Athlete == "Lowest Average" | Athlete == "XC Average" | Athlete == "Track Average")

averages_high <- radarCharting %>% 
  filter(Athlete == "Best Average" | Athlete == "XC Average" | Athlete == "Track Average")

averages_compared <- radarCharting %>% 
  filter(Athlete == "Best Average" | Athlete == "Lowest Average")

averages_teams <- radarCharting %>% 
  filter(Athlete == "Track Average" | Athlete == "XC Average")

# Extract numeric data for radar chart
radar_data <- averages[, numeric_cols]
best_radar_data <- averages_high[, numeric_cols]
comparison_radar_data <- averages_compared[, numeric_cols]
team_radar_data <- averages_teams[, numeric_cols]

# Add max and min rows for proper scaling
max_min <- data.frame(
  wgt_RSI = c(1, 0),
  wgt_Peak_Power = c(1, 0),
  wgt_Jump_Height = c(1, 0),
  wgt_Ecc_Peak_Force = c(1, 0),
  wgt_Con_Peak_Force = c(1, 0))

radar_data <- rbind(max_min, radar_data)
best_radar_data <- rbind(max_min, best_radar_data)
comparison_radar_data <- rbind(max_min, comparison_radar_data)
team_radar_data <- rbind(max_min, team_radar_data)

# Labels for each category (variables)
labels <- c("RSI-Mod", "Peak Power", "Jump Height", "Ecc Peak Force", "Con Peak Force")



# Create radar chart
radarchart(radar_data, axistype = 1, 
           pcol = c("#FF6347", "#4682B4", "#32CD32"), # Colors for the lines
           plwd = c(2.5, 1.5, 1.5), # Line widths
           plty = c(1, 2, 2), # Line types (1: solid, 2: dashed, 3: dotted)
           vlcex = .85, vlabels = labels, # Axis labels
           cglcol = "gray", cglty = 1, cglwd = 0.8, # Grid line customization
           axislabcol = "blue", # Axis label color
           caxislabels = seq(0, 1, 0.2), # Custom axis labels
           calcex = 0.75, # Axis label size
           pfcol = NA) # No polygon fill

# Add legend
legend("topright", legend = c("Lowest Average", "Track Average", "XC Average"), 
       col = c("#FF6347", "#4682B4", "#32CD32"), 
       lty = c(1, 2, 2), lwd = 2, 
       bty = "n")
title(main = "Lowest Timed Athletes Comparison to Team Averages")

par(mar = c(1.5, 1, 1, 1))
# Create radar chart
radarchart(best_radar_data, axistype = 1, 
           pcol = c("#32CD32", "#FF6347", "#4682B4"), # Colors for the lines
           plwd = c(2.5, 1.5, 1.5), # Line widths
           plty = c(1, 2, 2), # Line types (1: solid, 2: dashed, 3: dotted)
           vlcex = 1.05, vlabels = labels, # Axis labels
           cglcol = "gray", cglty = 1, cglwd = 0.8, # Grid line customization
           axislabcol = "blue", # Axis label color
           caxislabels = seq(0, 1, 0.2), # Custom axis labels
           calcex = 0.75,# Axis label size
            pfcol = c(rgb(50, 205, 50, maxColorValue = 255, alpha = 35),
                      rgb(70, 130, 180, maxColorValue = 255, alpha = 35), # Semi-transparent fill for XC Average
                      rgb(255, 99, 71, maxColorValue = 255, alpha = 35)))

# Add legend
legend("topright", legend = c("Best Average", "Track Average", "XC Average"), 
       col = c("#32CD32", "#FF6347", "#4682B4"), 
       lty = c(1, 2, 2), lwd = 2, 
       bty = "n")
title(main = "Best Timed Athletes Comparison to Team Averages")

par(mar = c(1.5, 1, 1, 1))
# Create radar chart
radarchart(comparison_radar_data, axistype = 1, 
           pcol = c("#4682B4", "#FF6347"), # Colors for the lines
           plwd = c(2.5, 2.5), # Line widths
           plty = c(2, 1), # Line types (1: solid, 2: dashed, 3: dotted)
           vlcex = 1.05, vlabels = labels, # Axis labels
           cglcol = "gray", cglty = 1, cglwd = 0.8, # Grid line customization
           axislabcol = "blue", # Axis label color
           caxislabels = seq(0, 1, 0.2), # Custom axis labels
           calcex = 0.75, # Axis label size
            pfcol = c(rgb(70, 130, 180, maxColorValue = 255, alpha = 85), # Semi-transparent fill for XC Average
                      rgb(255, 99, 71, maxColorValue = 255, alpha = 85)))

# Add legend
legend("topright", legend = c("Best Average", "Lowest Average"), 
       col = c("#FF6347", "#4682B4"), 
       lty = c(1, 2), lwd = 2, 
       bty = "n")
title(main = "Best vs. Worst Times Comparison")


par(mar = c(1.5, 1, 1, 1))
# Create radar chart
radarchart(team_radar_data, axistype = 1, 
           pcol = c( "#FF6347", "#4682B4"), # Colors for the lines
           plwd = c(2.5, 2.5), # Line widths
           plty = c(2, 1), # Line types (1: solid, 2: dashed, 3: dotted)
           vlcex = 1.05, vlabels = labels, # Axis labels
           cglcol = "gray", cglty = 1, cglwd = 0.8, # Grid line customization
           axislabcol = "blue", # Axis label color
           caxislabels = seq(0, 1, 0.2), # Custom axis labels
           calcex = 0.75, # Axis label size
           pfcol = c(rgb(255, 99, 71, maxColorValue = 255, alpha = 85),
                     rgb(70, 130, 180, maxColorValue = 255, alpha = 85) # Semi-transparent fill for XC Average
                     ))

# Add legend
legend("topright", legend = c("Track Average", "XC Average"), 
       col = c( "#FF6347", "#4682B4"), 
       lty = c(2, 1), lwd = 2, 
       bty = "n")
title(main = "Track vs. XC Times Comparison (No Athlete 1)")
```



